Comparing human and model-based forecasts of COVID-19 in Germany and Poland

### 2021-10-14 manuscript/manuscript.tex
clean code and files

We used two models from the \texttt{EpiNow2} R package (version 1.3.3) as our model-based forecasts (Abbott, Hellewell, et al. 2020). The first of these models, here called ``renewal model,'' used the renewal equation (Fraser 2007) to predict reported cases and deaths. It estimated the effective reproduction number \(R_t\) (the average number of people each infected person is expected to infect in turn) and models future infections as a weighted sum of past infections times \(R_t\). \(R_t\) was assumed to stay constant in the future, roughly corresponding to continuing the latest trend in infections. The \(R_t\) value which was carried forward was slightly changed on November 9th 2020, where before that date the last `reliable' \(R_t\) estimate from X days before the forecast was used and after that the latest and most up-to-date (but possibly less stable) \(R_t\) estimate made on the actual forecast date was used. Reported cases and deaths were obtained by convolving predicted infections over data-based delay distributions (Abbott, Hellewell, et al. 2020; DOI n.d.; {``Evaluating the Use of the Reproduction Number as an Epidemiological Tool, Using Spatio-Temporal Trends of the {Covid}-19 Outbreak in {England} \textbar{} {medRxiv}''} n.d.) to model the time between infection and report date or death. The renewal model was used to predict cases as well as deaths, but the two were forecast completely independently, meaning that \(R_t\) was forecast once from reported cases and once from reported deaths. Death forecasts from the renewal model were therefore not informed by past cases. One submission of the renewal model on December 28th 2020 was belated and therefore not included in the official Forecast hub ensemble.

The second model (``convolution model'') was only used to forecast deaths and was added later, starting December 7th 2020 (with the first forecast from December 7th suffering from a software bug and therefore disregarded in all further analyses). The convolution model was submitted, but never included in the official Forecast hub ensemble due to concerns that it could be too similar to the renewal model. The convolution model predicted deaths these as a fraction of infected people who would die with some delay, by using a convolution of reported cases with a distribution that described the delay from case report to death and a scaling factor (the case-fatality ratio). Both the renewal and the convolution model used daily observations and assumed a negative binomial observation model and a multiplicative day-of-the-week effect (Abbott, Hellewell, et al. 2020). Line list data used to inform the delay from symptom onset to test positive case report or death in the model-based forecasts was sourced from (Xu et al., n.d.) with data available up to the 1st of August. All model fitting was done using Markov-chain Monte Carlo (MCMC) in stan (Stan Development Team 2020) with each location and forecast target being fit separately. More details are available in the supplementary information.

@@

For the main analysis we focused on two week ahead predictions, as predictions beyond this horizon are often unreliable due to rapidly changing conditions (Bracher, Wolffram, et al. 2021). Forecasts for cases were scored using the full period from October 2020 until March 2021. To ensure comparability between models, all death forecasts were scored using only the period from December 14th on, where all models including the convolution model were available. To ensure robustness of our results we conducted a sensitivity analysis where all forecasts (including cases) where scored only over the later period where are forecasts were available (see Section \ref{sensitivity-analysis} in the SI). Results remained broadly unchanged.

@@

\begin{align}
  I_{t} &= I_0 \exp  \left(r t \right)  \\
  I_0 &\sim \mathcal{LN}(\log I_{obs}, 0.2) \\
  r &\sim \mathcal{LN}(r_{obs}, 0.2) 
\end{align}

Where \(I_{obs}\) and \(r_{obs}\) are estimated form the first week of observed data. For the time window of the observed data infections were then modelled by weighting previous infections by the generation time and scaling by the instantaneous reproduction number. These infections were then convolved to cases by date (\(O_t\)) and cases by date of report (\(D_t\)) using log-normal delay distributions. This model can be defined mathematically as follows,

\begin{align}
  \log R_{t} &= \log R_{t-1} + \mathrm{GP}_t \\
  I_t &= R_t \sum_{\tau = 1}^{15} w(\tau | \mu_{w}, \sigma_{w}) I_{t - \tau} \\
  O_t &= \sum_{\tau = 0}^{15} \xi_{O}(\tau | \mu_{\xi_{O}}, \sigma_{\xi_{O}}) I_{t-\tau} \\
  D_t &= \alpha \sum_{\tau = 0}^{15} \xi_{D}(\tau | \mu_{\xi_{D}}, \sigma_{\xi_{D}}) O_{t-\tau} \\ 
  C_t &\sim \mathrm{NB}\left(\omega_{(t \mod 7)}D_t, \phi\right)
\end{align}

Where,
\begin{align}
     w &\sim \mathcal{G}(\mu_{w}, \sigma_{w}) \\
    \xi_{O} &\sim \mathcal{LN}(\mu_{\xi_{O}}, \sigma_{\xi_{O}}) \\
    \xi_{D} &\sim \mathcal{LN}(\mu_{\xi_{D}}, \sigma_{\xi_{D}}) 
\end{align}

@@

The prior for the generation time was sourced from (Ganyani et al. 2020) but refit using a log-normal incubation period with a mean of 5.2 days (SD 1.1) and SD of 1.52 days (SD 1.1) with this incubation period also being used as a prior (Lauer et al. 2020) for \(\xi_{O}\). This resulted in a gamma-distributed generation time with mean 3.6 days (standard deviation (SD) 0.7), and SD of 3.1 days (SD 0.8) for all estimates. We estimated the delay between symyptom onset and case report or death required to convolve latent infections to observations by fitting an integer adjusted log-normal distribution to 10 subsampled bootstraps of a public linelist for cases in Germany from April 2020 to June 2020 with each bootstrap using 1\% or 1769 samples of the available data (Xu et al., n.d.; Abbott, Sherratt, et al. 2020) and combining the posteriors for the mean and standard deviation of the log-normal distribution (Abbott, Hellewell, et al. 2020; DOI n.d.; {``Evaluating the Use of the Reproduction Number as an Epidemiological Tool, Using Spatio-Temporal Trends of the {Covid}-19 Outbreak in {England} \textbar{} {medRxiv}''} n.d.; Stan Development Team 2020).

\(GP_t\) is an approximate Hilbert space Gaussian process as defined in (Riutort-Mayol et al. 2020) using a Matern 3/2 kernel using a boundary factor of 1.5 and 17 basis functions (20\% of the number of days used in fitting). The lengthscale of the Gaussian process was given a log-normal prior with a mean of 21 days, and a standard deviation of 7 days truncated to be greater than 3 days and less than 60 days. The magnitude of the Gaussian process was assumed be normally distributed centred at 0 with a standard deviation of 0.1.

@@

where \(I^c_t = \sum_{s< t} I_s\) are cumulative infections by \(t-1\) and \(I'_t\) are the unadjusted infections defined above. This adjustment is based on that implemented in the \texttt{epidemia} R package (Scott et al. 2020; \textbf{bhatt202?}).

@@

Both models were implemented using the \texttt{EpiNow2} R package (version 1.3.3) (Abbott, Hellewell, et al. 2020). Each forecast target was fit independently for each model using Markov-chain Monte Carlo (MCMC) in stan (Stan Development Team 2020). A minimum of 4 chains were used with a warmup of 250 samples for the renewal equation-based model and 1000 samples for the convolution model. 2000 samples total post warmup were used for the renewal equation model and 4000 samples for the convolution model. Different settings were chosen for each model to optimise compute time contigent on convergence. Convergence was assessed using the R hat diagnostic (Stan Development Team 2020). For the convolution model forecast the case forecast from the renewal equation model was used in place of observed cases beyond the forecast horizon using 1000 posterior samples. 12 weeks of data was used for both models though only 3 weeks of data were included in the likelihood for the convolution model.

@@

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{>{}llccccccccc}
\toprule
  & Model & WIS & WIS - sd & Dispersion & Underpred. & Overpred. & Bias & Abs. error & 50\%-Cov. & 90\%-Cov.\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Cases}}\\
\hline
\cellcolor{white}{} & Crowd forecast & 7010 (0.8) & 7480 (0.64) & 2680 (0.73) & 1700 (1.38) & 2630 (0.68) & -0.01 & 10400 (0.82) & 0.55 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 8770 (1) & 11700 (1) & 3670 (1) & 1230 (1) & 3870 (1) & -0.04 & 12700 (1) & 0.57 & 0.81\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 1 wk ahead}} & Renewal & 8740 (1) & 11800 (1.01) & 2190 (0.6) & 2720 (2.21) & 3830 (0.99) & 0.18 & 12000 (0.94) & 0.48 & 0.71\\
\cmidrule{1-11}
\cellcolor{white}{} & Crowd forecast & 16200 (0.89) & 16600 (0.76) & 3660 (0.6) & 5930 (1.56) & 6600 (0.78) & -0.01 & 23300 (0.87) & 0.36 & 0.55\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 18300 (1) & 21900 (1) & 6140 (1) & 3800 (1) & 8410 (1) & -0.03 & 26800 (1) & 0.43 & 0.64\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 2 wk ahead}} & Renewal & 25600 (1.4) & 33800 (1.54) & 5420 (0.88) & 5920 (1.56) & 14200 (1.69) & 0.17 & 34600 (1.29) & 0.43 & 0.67\\
\cmidrule{1-11}
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Deaths}}\\
\hline
\cellcolor{white}{} & Convolution & 255 (1.03) & 343 (1.01) & 82 (0.89) & 142 (1.23) & 31.1 (0.75) & -0.18 & 399 (1.19) & 0.42 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 265 (1.07) & 317 (0.94) & 78.2 (0.85) & 82 (0.71) & 105 (2.52) & 0.08 & 402 (1.2) & 0.38 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 248 (1) & 338 (1) & 92.2 (1) & 115 (1) & 41.6 (1) & -0.04 & 334 (1) & 0.62 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 1 wk ahead}} & Renewal & 298 (1.2) & 403 (1.19) & 87 (0.94) & 107 (0.93) & 105 (2.52) & -0.07 & 413 (1.24) & 0.50 & 0.79\\
\cmidrule{1-11}
\cellcolor{white}{} & Convolution & 357 (1.22) & 573 (1.49) & 104 (0.79) & 204 (1.89) & 48.8 (0.94) & -0.10 & 565 (1.32) & 0.33 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 368 (1.26) & 442 (1.15) & 107 (0.81) & 102 (0.94) & 160 (3.08) & 0.14 & 576 (1.34) & 0.38 & 0.75\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 292 (1) & 385 (1) & 132 (1) & 108 (1) & 51.9 (1) & 0.01 & 429 (1) & 0.62 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 2 wk ahead}} & Renewal & 524 (1.79) & 671 (1.74) & 155 (1.17) & 133 (1.23) & 236 (4.55) & -0.02 & 750 (1.75) & 0.50 & 0.71\\
\bottomrule
\end{tabular}}
\end{table}

@@

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{>{}llccccccccc}
\toprule
  & Model & WIS & WIS - sd & Dispersion & Underpred. & Overpred. & Bias & Abs. error & 50\%-Cov. & 90\%-Cov.\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Cases}}\\
\hline
\cellcolor{white}{} & Crowd forecast & 27000 (0.81) & 26200 (0.64) & 4750 (0.52) & 11000 (1.43) & 11200 (0.67) & 0.02 & 39000 (0.83) & 0.14 & 0.48\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 33400 (1) & 40700 (1) & 9130 (1) & 7690 (1) & 16600 (1) & -0.01 & 46900 (1) & 0.29 & 0.62\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 3 wk ahead}} & Renewal & 50600 (1.51) & 70000 (1.72) & 10800 (1.18) & 7710 (1) & 32100 (1.93) & 0.13 & 68700 (1.46) & 0.29 & 0.55\\
\cmidrule{1-11}
\cellcolor{white}{} & Crowd forecast & 39200 (0.7) & 38600 (0.52) & 5970 (0.49) & 15600 (1.26) & 17600 (0.56) & 0.07 & 54800 (0.74) & 0.05 & 0.38\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 55900 (1) & 73700 (1) & 12200 (1) & 12400 (1) & 31300 (1) & 0.01 & 74400 (1) & 0.24 & 0.52\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 4 wk ahead}} & Renewal & 91700 (1.64) & 135000 (1.83) & 19500 (1.6) & 8990 (0.72) & 63200 (2.02) & 0.09 & 125000 (1.68) & 0.31 & 0.48\\
\cmidrule{1-11}
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Deaths}}\\
\hline
\cellcolor{white}{} & Convolution & 541 (1.7) & 802 (2.45) & 157 (0.91) & 279 (3.01) & 105 (1.91) & -0.04 & 747 (1.53) & 0.54 & 0.75\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 414 (1.3) & 526 (1.6) & 137 (0.8) & 82 (0.88) & 194 (3.52) & 0.12 & 648 (1.33) & 0.42 & 0.83\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 319 (1) & 328 (1) & 172 (1) & 92.7 (1) & 55.1 (1) & -0.03 & 488 (1) & 0.54 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 3 wk ahead}} & Renewal & 724 (2.27) & 916 (2.79) & 249 (1.45) & 158 (1.7) & 317 (5.75) & -0.01 & 1040 (2.13) & 0.46 & 0.83\\
\cmidrule{1-11}
\cellcolor{white}{} & Convolution & 763 (1.8) & 932 (2.1) & 268 (1.26) & 331 (2.63) & 164 (1.91) & 0.01 & 985 (1.46) & 0.54 & 0.75\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 498 (1.17) & 633 (1.43) & 168 (0.79) & 83.6 (0.66) & 246 (2.87) & 0.14 & 756 (1.12) & 0.38 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 424 (1) & 443 (1) & 212 (1) & 126 (1) & 85.7 (1) & -0.06 & 675 (1) & 0.58 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 4 wk ahead}} & Renewal & 959 (2.26) & 1210 (2.73) & 337 (1.59) & 200 (1.59) & 421 (4.91) & -0.05 & 1350 (2) & 0.50 & 0.79\\
\bottomrule
\end{tabular}}
\end{table}

@@

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/aggregate-performance-2-weeks-locations-all-4} \caption{Visualisation of aggregate performance metrics across locations. A: mean weighted interval score (WIS) across horizons. B: median WIS. C: Absolute error of the median forecast. D: Standard deviation of the WIS. E: Dispersion (higher values mean further spread out forecast). F: Bias, i.e. general tendency to over- or underpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Empirical coverage of the 50\% prediction intervals. F: Empirical coverage of the 90\% prediction intervals.}\label{fig:performance-locations}
\end{figure}

@@

>>>>

We used two models from the \texttt{EpiNow2} R package (version 1.3.3) as our model-based forecasts (Abbott, Hellewell, et al. 2020). The first of these models, here called ``renewal model,'' used the renewal equation (Fraser 2007) to predict reported cases and deaths. It estimated the effective reproduction number \(R_t\) (the average number of people each infected person is expected to infect in turn) and models future infections as a weighted sum of past infection multiplied by \(R_t\). \(R_t\) was assumed to stay constant beyond the forecast horizon, roughly corresponding to continuing the latest exponential trend in infections. On the 9th of November we altered the date when \(R_t\) was assumed to be constant from two weeks prior to the date of the forecast to the forecast date. Reported notifications were obtained by convolving predicted infections over data-based delay distributions (Abbott, Hellewell, et al. 2020; DOI n.d.; {``Evaluating the Use of the Reproduction Number as an Epidemiological Tool, Using Spatio-Temporal Trends of the {Covid}-19 Outbreak in {England} \textbar{} {medRxiv}''} n.d.) to model the time between infection and report date or death. The renewal model was used to predict cases as well as deaths with forecasts being generated for each target separately, meaning that \(R_t\) was forecast once from reported cases and once from reported deaths. Death forecasts from the renewal model were therefore not informed by past cases. One submission of the renewal model on December 28th 2020 was delayed and therefore not included in the official Forecast hub ensemble.

The second model (``convolution model'') was only used to forecast deaths and was added later, starting December 7th 2020 (with the first forecast from December 7th suffering from a software bug and therefore disregarded in all further analyses). The convolution model was submitted, but never included in the official Forecast hub ensemble due to concerns that it could be too similar to the renewal model. The convolution model predicted deaths as a fraction of infected people who would die with some delay, by using a convolution of reported cases with a distribution that described the delay from case report to death and a scaling factor (the case-fatality ratio). Both the renewal and the convolution model used daily observations and assumed a negative binomial observation model with a multiplicative day-of-the-week effect (Abbott, Hellewell, et al. 2020). Line list data used to inform the prior for the delay from symptom onset to test positive case report or death in the model-based forecasts was sourced from (Xu et al., n.d.) with data available up to the 1st of August. All model fitting was done using Markov-chain Monte Carlo (MCMC) in stan (Stan Development Team 2020) with each location and forecast target being fit separately. More details are available in the supplementary information.

@@

For the main analysis we focused on two week ahead predictions, as predictions beyond this horizon are often unreliable due to rapidly changing conditions (Bracher, Wolffram, et al. 2021). Forecasts for cases were scored using the full period from October 2020 until March 2021. To ensure comparability between models, all death forecasts were scored using only the period from December 14th on, where all models including the convolution model were available. To ensure robustness of our results we conducted a sensitivity analysis where all forecasts (including cases) were scored only over the later period where forecasts were available (see Section \ref{sensitivity-analysis} in the SI). Results remained broadly unchanged.

@@

\begin{align*}
  I_{t} &= I_0 \exp  \left(r t \right)  \\
  I_0 &\sim \mathcal{LN}(\log I_{obs}, 0.2) \\
  r &\sim \mathcal{LN}(r_{obs}, 0.2) 
\end{align*}

Where \(I_{obs}\) and \(r_{obs}\) are estimated from the first week of observed data. For the time window of the observed data infections were then modelled by weighting previous infections by the generation time and scaling by the instantaneous reproduction number. These infections were then convolved to cases by date (\(O_t\)) and cases by date of report (\(D_t\)) using log-normal delay distributions. This model can be defined mathematically as follows,

\begin{align*}
  \log R_{t} &= \log R_{t-1} + \mathrm{GP}_t \\
  I_t &= R_t \sum_{\tau = 1}^{15} w(\tau | \mu_{w}, \sigma_{w}) I_{t - \tau} \\
  O_t &= \sum_{\tau = 0}^{15} \xi_{O}(\tau | \mu_{\xi_{O}}, \sigma_{\xi_{O}}) I_{t-\tau} \\
  D_t &= \alpha \sum_{\tau = 0}^{15} \xi_{D}(\tau | \mu_{\xi_{D}}, \sigma_{\xi_{D}}) O_{t-\tau} \\ 
  C_t &\sim \mathrm{NB}\left(\omega_{(t \mod 7)}D_t, \phi\right)
\end{align*}

Where,
\begin{align*}
     w &\sim \mathcal{G}(\mu_{w}, \sigma_{w}) \\
    \xi_{O} &\sim \mathcal{LN}(\mu_{\xi_{O}}, \sigma_{\xi_{O}}) \\
    \xi_{D} &\sim \mathcal{LN}(\mu_{\xi_{D}}, \sigma_{\xi_{D}}) 
\end{align*}

@@

The prior for the generation time was sourced from (Ganyani et al. 2020) but refit using a log-normal incubation period with a mean of 5.2 days (SD 1.1) and SD of 1.52 days (SD 1.1) with this incubation period also being used as a prior (Lauer et al. 2020) for \(\xi_{O}\). This resulted in a gamma-distributed generation time with mean 3.6 days (standard deviation (SD) 0.7), and SD of 3.1 days (SD 0.8) for all estimates. We estimated the delay between symptom onset and case report or death required to convolve latent infections to observations by fitting an integer adjusted log-normal distribution to 10 subsampled bootstraps of a public linelist for cases in Germany from April 2020 to June 2020 with each bootstrap using 1\% or 1769 samples of the available data (Xu et al., n.d.; Abbott, Sherratt, et al. 2020) and combining the posteriors for the mean and standard deviation of the log-normal distribution (Abbott, Hellewell, et al. 2020; DOI n.d.; {``Evaluating the Use of the Reproduction Number as an Epidemiological Tool, Using Spatio-Temporal Trends of the {Covid}-19 Outbreak in {England} \textbar{} {medRxiv}''} n.d.; Stan Development Team 2020).

\(GP_t\) is an approximate Hilbert space Gaussian process as defined in (Riutort-Mayol et al. 2020) using a Matern 3/2 kernel using a boundary factor of 1.5 and 17 basis functions (20\% of the number of days used in fitting). The length scale of the Gaussian process was given a log-normal prior with a mean of 21 days, and a standard deviation of 7 days truncated to be greater than 3 days and less than 60 days. The magnitude of the Gaussian process was assumed to be normally distributed centred at 0 with a standard deviation of 0.1.

@@

where \(I^c_t = \sum_{s< t} I_s\) are cumulative infections by \(t-1\) and \(I'_t\) are the unadjusted infections defined above. This adjustment is based on that implemented in the \texttt{epidemia} R package (Scott et al. 2020; Bhatt et al., n.d.).

@@

Both models were implemented using the \texttt{EpiNow2} R package (version 1.3.3) (Abbott, Hellewell, et al. 2020). Each forecast target was fitted independently for each model using Markov-chain Monte Carlo (MCMC) in stan (Stan Development Team 2020). A minimum of 4 chains were used with a warmup of 250 samples for the renewal equation-based model and 1000 samples for the convolution model. 2000 samples total post warmup were used for the renewal equation model and 4000 samples for the convolution model. Different settings were chosen for each model to optimise compute time contingent on convergence. Convergence was assessed using the R hat diagnostic (Stan Development Team 2020). For the convolution model forecast the case forecast from the renewal equation model was used in place of observed cases beyond the forecast horizon using 1000 posterior samples. 12 weeks of data was used for both models though only 3 weeks of data were included in the likelihood for the convolution model.

@@

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{>{}llccccccccc}
\toprule
  & Model & WIS & WIS - sd & dispersion & Underpred. & Overpred. & Bias & Abs. error & 50\%-Cov. & 90\%-Cov.\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Cases}}\\
\hline
\cellcolor{white}{} & Crowd forecast & 7010 (0.8) & 7480 (0.64) & 2680 (0.73) & 1700 (1.38) & 2630 (0.68) & -0.01 & 10400 (0.82) & 0.55 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 8770 (1) & 11700 (1) & 3670 (1) & 1230 (1) & 3870 (1) & -0.04 & 12700 (1) & 0.57 & 0.81\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 1 wk ahead}} & Renewal & 8740 (1) & 11800 (1.01) & 2190 (0.6) & 2720 (2.21) & 3830 (0.99) & 0.18 & 12000 (0.94) & 0.48 & 0.71\\
\cmidrule{1-11}
\cellcolor{white}{} & Crowd forecast & 16200 (0.89) & 16600 (0.76) & 3660 (0.6) & 5930 (1.56) & 6600 (0.78) & -0.01 & 23300 (0.87) & 0.36 & 0.55\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 18300 (1) & 21900 (1) & 6140 (1) & 3800 (1) & 8410 (1) & -0.03 & 26800 (1) & 0.43 & 0.64\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 2 wk ahead}} & Renewal & 25600 (1.4) & 33800 (1.54) & 5420 (0.88) & 5920 (1.56) & 14200 (1.69) & 0.17 & 34600 (1.29) & 0.43 & 0.67\\
\cmidrule{1-11}
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Deaths}}\\
\hline
\cellcolor{white}{} & Convolution & 255 (1.03) & 343 (1.01) & 82 (0.89) & 142 (1.23) & 31.1 (0.75) & -0.18 & 399 (1.19) & 0.42 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 265 (1.07) & 317 (0.94) & 78.2 (0.85) & 82 (0.71) & 105 (2.52) & 0.08 & 402 (1.2) & 0.38 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 248 (1) & 338 (1) & 92.2 (1) & 115 (1) & 41.6 (1) & -0.04 & 334 (1) & 0.62 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 1 wk ahead}} & Renewal & 298 (1.2) & 403 (1.19) & 87 (0.94) & 107 (0.93) & 105 (2.52) & -0.07 & 413 (1.24) & 0.50 & 0.79\\
\cmidrule{1-11}
\cellcolor{white}{} & Convolution & 357 (1.22) & 573 (1.49) & 104 (0.79) & 204 (1.89) & 48.8 (0.94) & -0.10 & 565 (1.32) & 0.33 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 368 (1.26) & 442 (1.15) & 107 (0.81) & 102 (0.94) & 160 (3.08) & 0.14 & 576 (1.34) & 0.38 & 0.75\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 292 (1) & 385 (1) & 132 (1) & 108 (1) & 51.9 (1) & 0.01 & 429 (1) & 0.62 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 2 wk ahead}} & Renewal & 524 (1.79) & 671 (1.74) & 155 (1.17) & 133 (1.23) & 236 (4.55) & -0.02 & 750 (1.75) & 0.50 & 0.71\\
\bottomrule
\end{tabular}}
\end{table}

@@

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{>{}llccccccccc}
\toprule
  & Model & WIS & WIS - sd & dispersion & Underpred. & Overpred. & Bias & Abs. error & 50\%-Cov. & 90\%-Cov.\\
\midrule
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Cases}}\\
\hline
\cellcolor{white}{} & Crowd forecast & 27000 (0.81) & 26200 (0.64) & 4750 (0.52) & 11000 (1.43) & 11200 (0.67) & 0.02 & 39000 (0.83) & 0.14 & 0.48\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 33400 (1) & 40700 (1) & 9130 (1) & 7690 (1) & 16600 (1) & -0.01 & 46900 (1) & 0.29 & 0.62\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 3 wk ahead}} & Renewal & 50600 (1.51) & 70000 (1.72) & 10800 (1.18) & 7710 (1) & 32100 (1.93) & 0.13 & 68700 (1.46) & 0.29 & 0.55\\
\cmidrule{1-11}
\cellcolor{white}{} & Crowd forecast & 39200 (0.7) & 38600 (0.52) & 5970 (0.49) & 15600 (1.26) & 17600 (0.56) & 0.07 & 54800 (0.74) & 0.05 & 0.38\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 55900 (1) & 73700 (1) & 12200 (1) & 12400 (1) & 31300 (1) & 0.01 & 74400 (1) & 0.24 & 0.52\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-3}{*}{\raggedright\arraybackslash 4 wk ahead}} & Renewal & 91700 (1.64) & 135000 (1.83) & 19500 (1.6) & 8990 (0.72) & 63200 (2.02) & 0.09 & 125000 (1.68) & 0.31 & 0.48\\
\cmidrule{1-11}
\addlinespace[0.3em]
\multicolumn{11}{l}{\textbf{Deaths}}\\
\hline
\cellcolor{white}{} & Convolution & 541 (1.7) & 802 (2.45) & 157 (0.91) & 279 (3.01) & 105 (1.91) & -0.04 & 747 (1.53) & 0.54 & 0.75\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 414 (1.3) & 526 (1.6) & 137 (0.8) & 82 (0.88) & 194 (3.52) & 0.12 & 648 (1.33) & 0.42 & 0.83\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 319 (1) & 328 (1) & 172 (1) & 92.7 (1) & 55.1 (1) & -0.03 & 488 (1) & 0.54 & 0.96\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 3 wk ahead}} & Renewal & 724 (2.27) & 916 (2.79) & 249 (1.45) & 158 (1.7) & 317 (5.75) & -0.01 & 1040 (2.13) & 0.46 & 0.83\\
\cmidrule{1-11}
\cellcolor{white}{} & Convolution & 763 (1.8) & 932 (2.1) & 268 (1.26) & 331 (2.63) & 164 (1.91) & 0.01 & 985 (1.46) & 0.54 & 0.75\\
\cmidrule{2-11}
\cellcolor{white}{} & Crowd forecast & 498 (1.17) & 633 (1.43) & 168 (0.79) & 83.6 (0.66) & 246 (2.87) & 0.14 & 756 (1.12) & 0.38 & 0.79\\
\cmidrule{2-11}
\cellcolor{white}{} & Hub-ensemble & 424 (1) & 443 (1) & 212 (1) & 126 (1) & 85.7 (1) & -0.06 & 675 (1) & 0.58 & 0.92\\
\cmidrule{2-11}
\cellcolor{white}{\multirow{-4}{*}{\raggedright\arraybackslash 4 wk ahead}} & Renewal & 959 (2.26) & 1210 (2.73) & 337 (1.59) & 200 (1.59) & 421 (4.91) & -0.05 & 1350 (2) & 0.50 & 0.79\\
\bottomrule
\end{tabular}}
\end{table}

@@

\begin{figure}[H]
\includegraphics[width=1\linewidth,]{../analysis/plots/aggregate-performance-2-weeks-locations-all-v4} \caption{Visualisation of aggregate performance metrics across locations. A: mean weighted interval score (WIS) across horizons. B: median WIS. C: Absolute error of the median forecast. D: Standard deviation of the WIS. E: Dispersion (higher values mean further spread out forecast). F: Bias, i.e. general tendency to over- or underpredict. Values are between -1 (complete under-prediction) and 1 (complete over-prediction) and 0 ideally. G: Empirical coverage of the 50\% prediction intervals. F: Empirical coverage of the 90\% prediction intervals.}\label{fig:performance-locations}
\end{figure}

@@

\leavevmode\vadjust pre{\hypertarget{ref-bhattSemiMechanisticBayesianModeling}{}}%
Bhatt, Samir, Neil Ferguson, Seth Flaxman, Axel Gandy, Swapnil Mishra, and James A Scott. n.d. {``Semi-{Mechanistic Bayesian} Modeling of {COVID}-19 with {Renewal Processes},''} 14.
